X-ray phase-contrast imaging at \SI{100}{\kilo\electronvolt} on table top sources

### 2013-10-23 hedpc_natphys.tex
Appeal edits
    - include table top sources in title
    - shorten introduction
    - more small cosmetic rephrasings

%acronyms
\acrodef{CT}{computed tomography}
\acrodef{SNR}{signal-to-noise ratio}
\acrodef{CNR}{contrast-to-noise ratio}
%% ----------------------------------------------------------------------------------------------------------------

@@

\title{X-ray phase-contrast imaging at \SI{100}{\kilo\electronvolt}}

@@

The attenuation and phase shift properties are described by the complex index of refraction
$n=1-\delta(\mathbf{r},\lambda) + i \beta(\mathbf{r},\lambda)$.
The imaginary part $\beta(\mathbf{r},\lambda)$ is related to the
attenuation per unit length (attenuation coefficient) by $\mu = 4 \pi
\beta(\mathbf{r},\lambda) / \lambda$ and the total attenuation by an object
of thickness $d$ can be described by the line integral $L_\mu = \int_0^d
\mu(\mathbf{r},\lambda) \d{} \mathbf{r}$. Similarly, the real part
$\delta(\mathbf{r},\lambda)$ determines the phase shift per unit length by
$\phi = 2 \pi \delta / \lambda$ and the total phase shift is $L_\phi =
\int_0^d \phi(\mathbf{r}) \d{} \mathbf{r}$.

While the attenuation line integral $L_\mu$ can be directly measured with an
X-ray detector by measuring the reduction of the beam intensity $I$ ($L_\mu
\propto \log(I)$), the measurement of the phase shift integral $L_\phi$ is
more challenging as there is no such simple relation. However, phase-sensitive
imaging is a desirable modality, as it can provide an enhanced
\ac{CNR} in images compared to attenuation for certain
materials or in different tissues \cite{Pfeiffer2007a,McDonald2009}.
Furthermore, the refractive index decrement $\delta$ provides direct access
to the electron density \cite{Als-Nielsen2011} and in combination with
attenuation, it further enables the determination of the effective atomic
number of a material \cite{Qi2010}.

In the past, a lot of effort has been invested into the development of
techniques which are sensitive to the phase shift. Since most of the
available phase contrast techniques rely on interference and thus typically
on optical hardware (e.g. crystals, gratings), they vary a lot in terms of
sensitivity, practical applicability or achievable resolution. The vast
majority of the methods, including crystal analyzer based
\cite{Davis1995,Chapman1997} or interferometric \cite{Bonse1965,Momose1996}
methods rely on X-ray beams of high spatial and temporal coherence, which is
available only at synchrotron sources. Techniques which are compatible with
an X-ray beam of low temporal coherence (broad bandwidth) are the inline
phase contrast method \cite{Snigirev1995,Wilkins1996,Cloetens1996} and
Talbot interferometry \cite{Cloetens1997,David2002,Momose2003a}.
Phase-contrast imaging using X-ray beams of low temporal \emph{and}
spatial coherence such as conventional low-brilliance X-ray tubes have been
demonstrated with Talbot-Lau interferometry \cite{Pfeiffer2006} and coded
apertures \cite{Munro2012}.
In addition to phase sensitivity, these techniques provide a dark-field
contrast, also called scatter contrast or visibility-reduction contrast,
that is proportional to the integrated local small angle scattering power
from microscopic density fluctuations in a specimen\cite{Pfeiffer2008}.

@@

Here, we introduce a method for phase-contrast imaging which is
compatible with the entire diagnostic energy range of X-rays. Moreover, the
method is fully compatible with compact imaging arrangements based on
conventional X-ray tubes. The method is based on Talbot-Lau interferometry
\cite{Pfeiffer2006} and employs an edge-on approach for the grating design and
arrangement. 
The challenge which limited the progress towards higher energies was mainly
related to the manufacturing of gratings with high aspect ratios. The aspect
ratio, given by
\begin{equation}
    \text{AR} = \frac{2h}{p},
\end{equation}
where $p$ is the grating period and $h$ the grating structure
height, is normally limited by the lithographic process, as grating
structures tend to collapse or to deform (e.g. through capillary forces) if
the aspect ratio is too high. For a given setup distance these parameters
depend on the target energy $E$ according to $p \propto
1/\sqrt{E}$ and $h \propto E^3$, and therefore $\textnormal{AR}
\propto E^{7/2}$\cite{Momose2003a}. If at $E=\SI{25}{\kilo\electronvolt}$ an aspect ratio
for the absorption grating of around $\textnormal{AR}=30$ is necessary for a
reasonable length of the experimental arrangement, it would have to be at
least $128$ for $E=\SI{100}{\kilo\electronvolt}$. Moreover, when using a
broad spectrum, photons above the design energy should also be
efficiently blocked by the gratings, which requires even higher aspect ratios.
Maximum achievable aspect ratios of current grating fabrication techniques
\cite{David2007,Kenntner2010} are approximately around 60,
larger values are often not achievable or come at the expense of a very poor
grating quality and performance.

Our design introduces the edge-on
illumination of circularly aligned structures. Edge-on illumination
(figure~\ref{Fig:schematic}), as
opposed to face-on illumination, exploits the dimension along the grating
lines to form a high aspect ratio of the structures in direction of the beam. The
effective structure height of the grating is then determined by the grating
dimension along the grating lines, which essentially allows arbitrarily high
aspect ratios. 
\begin{figure}[ht]
    \includegraphics[width=.7\textwidth]{figures/figure1.eps}
    \caption{Schematic of a grating
        interferometer for X-ray energies between 60 and
        \SI{150}{\kilo\electronvolt} in edge-on illumination mode. The
        aspect ratio is defined by the ratio of the travelling distance along the
        grating lines and the period and can be arbitrarily long. In order to avoid
        a reduction of the field of view, the grating structures are aligned on an
    arc.}
    \label{Fig:schematic}
\end{figure}

Increasing the aspect ratio of the gratings typically leads to a
reduction of the field of view due to the change of the grating transmission
function at high incident angles, which has also been identified by using a
glancing angle of the gratings between zero and \SI{90}{\degree}
\cite{Stutman2012a}. In order to overcome this problem, the grating lines are circularly aligned 
with a radius equal to the distance to the source.

@@

Grating design and fabrication is nonstandard and involves a complex mask
design, as shown in Figure~\ref{Fig:grating_mask}. Each grating resides on a
silicon chip and has its specific structure length and curvature. For the
current experiments, a symmetric interferometer with a grating period of $p
= \SI{2.8}{\micro\metre}$ for all gratings has been used. The design energy
is $\SI{100}{\kilo\electronvolt}$ and the beam splitter grating periodically
shifts the phase by zero and $\pi$ at this energy \cite{David2002}. Using
gold as the phase shifting material, this requires a structure length of
$h_1 = \SI{19.8}{\micro \metre}$. The analyzer grating is an absorption mask
for sensing slight changes of the interference pattern generated by the beam
splitter \cite{Momose2003a}. With a structure length of $h_2 =
\SI{800}{\micro \metre}$, this grating has an aspect ratio of $2h/p \approx
570$ and thus sufficiently attenuates X-rays up to energies of 
$\SI{160}{\kilo\electronvolt}$. Beam splitter- and analyzer grating are
separated at the first fractional Talbot order \cite{Weitkamp2005},
resulting in an intergrating distance of $\SI{158}{\milli\metre}$. The
source grating splits the relatively large focal spot ($\sim
\SI{1}{\milli\metre}$) into an array of individually coherent, but mutually
incoherent sources \cite{Pfeiffer2006}. It is also made of gold structures
with a structure length of $h_0 = h_2 = \SI{800}{\micro \metre}$.
\begin{figure} [ht]
    \includegraphics[width=.6\textwidth]{figures/grating_mask.eps}
    \caption{Grating design mask for
        the edge-on illumination approach and scanning electron microscope
        image of the grating. The top part of the 4 inch wafer shows
        five grating chips. The
        gratings have different curvatures which are specific to the grating
        interferometer geometry. 
        The SEM image shows the gold structures and the interrupting bridges
        that prevent the lamellae from collapsing\cite{Kenntner2010}.}
        \label{Fig:grating_mask}
\end{figure}

Due to the high spectral acceptance \cite{Weitkamp2005,Thuering2013c} of the
interferometer ($\SI{50}{\kilo\electronvolt}$ to
more than $\SI{160}{\kilo\electronvolt}$) and the high attenuation efficiencies of
the source and analyzer gratings ($>90\%$ up to
$\SI{160}{\kilo\electronvolt}$), the voltage of the X-ray source was set to
the maximum of $\SI{160}{\kilo\volt}$. With a grating structure height of
approx. $\SI{100}{\micro \metre}$, the field of view in the vertical
direction is limited to one detector pixel row. In the horizontal direction,
the field of view is limited by the grating size and the geometric
magnification of the sample, yielding a maximum field of view of
$\SI{30}{\milli\metre}$. In addition to the standard components (source,
camera, interferometer), two optical slits, one in front of the source
grating, the other in front of the camera, were required for the collimation
of the beam in the vertical direction. X-rays which are not travelling
through all of the gratings do not contribute to the signal and are
attenuated by the slits.

Fig.~\ref{Fig:img_chip} shows a radiographic scan of an electronic chip.
Several resistors and an integrated circuits are located on different layers
on the chip. The images were acquired in scanning mode, using a step size of
$\SI{100}{\micro \metre}$ along the $y$ axis. For a better comparison of the
magnified phase and attenuation images, the attenuation image has been
replaced with the differential attenuation image, which was obtained by
digital differentiation. In the attenuation image, the contrast of the soldering points of
the integrated circuit is reduced underneath the resistors, while in
the phase image, they can clearly be identified. This shows the
benefit of the phase nature of the image, which may be
useful to identify flaws in multilayered structures such as electronic
chips.
\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{figures/img_chip_dabs.eps}
    \caption{Radiographic scan of an electronic chip. The image was acquired
        with 24 phase steps per line and an exposure time of \SI{15}{\second} per
    step. The top right image shows the differential absorption image.}
    \label{Fig:img_chip}
\end{figure}

Edge-on illuminated grating interferometry breaks the current limitations of
phase-contrast imaging to the lower diagnostic energy range. Compact
geometries  using conventional X-ray sources for design energies well above
$\SI{100}{\kilo\electronvolt}$ could be realized with the approach. This
enables the examination of materials of higher density or thickness which
would be intransparent at lower energies. Finally, the approach is not
limited to X-ray imaging, but could easily be applied to other grating-based
imaging modalities (e.g., with neutrons \cite{Grunzweig2008}) where high
aspect ratios are required.

@@

In Fig.~\ref{Fig:img_chip}, image acquisition involved 24 phase steps per
line \cite{Weitkamp2005} and an exposure time of 15 seconds per step. The long
exposure times are mostly constrained by the low average visibility of the
gratings (\SI{5}{\percent}). The exposure time can be shortened with better
gratings and an efficient detector.

@@

>>>>

%acronyms
\acrodef{CT}{computed tomography}
\acrodef{SNR}{signal-to-noise ratio}
\acrodef{CNR}{contrast-to-noise ratio}
\acrodef{SEM}{scanning electron microscope}
%% ----------------------------------------------------------------------------------------------------------------

@@

\title{X-ray phase-contrast imaging at \SI{100}{\kilo\electronvolt} on
table top sources}

@@

The attenuation and phase shift properties are described by the complex index of refraction
$n=1-\delta + i \beta$.
The imaginary part $\beta$ is related to the
attenuation coefficient by $\mu = 4 \pi
\beta(\lambda) / \lambda$, while the real part
$\delta(\lambda)$ determines the phase shift 
$\phi = 2 \pi \delta(\lambda) / \lambda$

While the attenuation can be measured with an
X-ray detector as the reduction of the beam intensity, the phase
is not directly observable. Therefore, an optical system is needed to
convert the phase shift into intensity modulations.

Phase-sensitive
imaging is a desirable modality, as it can provide a complementary source of
contrast with respect to absorption by providing direct access 
to the electron density \cite{Als-Nielsen2011}. Moreover, the combination with
attenuation enables the determination of the effective atomic
number \cite{Qi2010}.
An enhanced
\ac{CNR} in images compared to attenuation for certain
materials or tissues \cite{Pfeiffer2007a,McDonald2009} has also been
demonstrated.

The vast majority of phase-sensitive techniques, including crystal analyzer based
\cite{Davis1995,Chapman1997} or interferometric \cite{Bonse1965,Momose1996}
methods rely on X-ray beams of high spatial and temporal coherence, which is
available only at synchrotron sources. Inline
phase contrast \cite{Snigirev1995,Wilkins1996,Cloetens1996} and
Talbot interferometry \cite{Cloetens1997,David2002,Momose2003a} need high
spatial coherence but are available on microfocus sources.
Phase-contrast imaging using X-ray beams of low temporal \emph{and}
spatial coherence such as conventional low-brilliance X-ray tubes have been
demonstrated with coded
apertures \cite{Munro2012} and Talbot-Lau interferometry \cite{Pfeiffer2006}.

WHAT IS TLINTERFEROMETER

In addition to phase sensitivity, these techniques provide a dark-field
contrast, also called scatter contrast or visibility-reduction contrast,
that is proportional to the integrated local small angle scattering power
from microscopic density fluctuations in a specimen\cite{Pfeiffer2008}.

@@

Here, we introduce a method for phase-contrast imaging which is
compatible with the entire diagnostic energy range of X-rays, 
compact imaging arrangements and 
conventional X-ray tubes. The method is based on Talbot-Lau interferometry
\cite{Pfeiffer2006} and employs an edge-on approach for the grating design and
arrangement. 
The challenge which limited the progress towards higher energies was mainly
related to the manufacturing of gratings with high aspect ratios. The aspect
ratio, given by
\begin{equation}
    \text{R} = \frac{2h}{p},
\end{equation}
where $p$ is the grating period and $h$ the  structure
height, is normally limited by the lithographic process, as grating
structures tend to collapse or deform (e.g. due to capillary forces) if
the aspect ratio is too high. For a given setup length these parameters
depend on the target energy $E$ according to $p \propto
1/\sqrt{E}$ and $h \propto E^3$, and therefore $\textnormal{R}
\propto E^{7/2}$~\cite{Momose2003a}. If at $E=\SI{25}{\kilo\electronvolt}$ an aspect ratio
for the absorption grating around $\textnormal{R}=30$ is necessary for a
reasonable length of the experimental arrangement, it would have to be at
least $128$ for $E=\SI{100}{\kilo\electronvolt}$. Moreover, when using a
broad spectrum, photons above the design energy should also be
efficiently blocked by the gratings, which requires even higher aspect ratios.
The maximum achievable aspect ratios of current fabrication techniques
\cite{David2007,Kenntner2010} are around 60.

Our design introduces the edge-on
illumination of circularly aligned structures. Edge-on illumination
(figure~\ref{Fig:schematic}), as
opposed to face-on illumination, exploits the dimension along the grating
lines to form a high aspect ratio of the structures in the direction of the beam. The
effective structure height of the grating is then determined by the grating
dimension along the grating lines, which essentially allows arbitrarily high
aspect ratios. 
\begin{figure}[ht]
    \includegraphics[width=.7\textwidth]{figures/figure1.eps}
    \caption{Schematic of a grating
        interferometer for X-ray energies between 60 and
        \SI{150}{\kilo\electronvolt} in edge-on illumination mode. The
        aspect ratio is defined by the ratio of the travelling distance along the
        grating lines and the period and can be arbitrarily long. In order to avoid
        a reduction of the field of view, the grating structures are aligned on an
    arc.}
    \label{Fig:schematic}
\end{figure}

Increasing the aspect ratio of the gratings typically leads to a
reduction of the field of view due to the change of the grating transmission
function at high incident angles. In order to overcome this problem, the grating lines are circularly aligned 
with a radius equal to the distance to the source.

@@

Grating design and fabrication is nonstandard and involves a complex mask
design, as shown in Figure~\ref{Fig:grating_mask}. Each grating resides on a
silicon chip and has its specific structure length and curvature. For the
current experiments, a symmetric interferometer with a grating period of $p
= \SI{2.8}{\micro\metre}$ for all gratings has been used. The design energy
is $\SI{100}{\kilo\electronvolt}$ and the beam splitter grating periodically
shifts the phase by zero and $\pi$ at this energy \cite{David2002}. Using
gold as the phase shifting material, this requires a structure length of
$h_1 = \SI{19.8}{\micro \metre}$. The analyzer grating is an absorption mask
for sensing slight changes of the interference pattern generated by the beam
splitter \cite{Momose2003a}. With a structure length of $h_2 =
\SI{800}{\micro \metre}$
it can sufficiently attenuate X-rays up to energies of 
$\SI{160}{\kilo\electronvolt}$. Beam splitter and analyzer grating are
separated at the first fractional Talbot order \cite{Weitkamp2005},
resulting in an intergrating distance of $\SI{158}{\milli\metre}$. The
source grating splits the relatively large focal spot ($\sim
\SI{1}{\milli\metre}$) into an array of individually coherent, but mutually
incoherent sources \cite{Pfeiffer2006}. It is also made of gold structures
with a length of $h_0 = h_2 = \SI{800}{\micro \metre}$.
\begin{figure} [ht]
    \includegraphics[width=.6\textwidth]{figures/grating_mask.eps}
    \caption{Grating design mask for
        the edge-on illumination approach and \ac{SEM}
        image of the grating. The top part of the 4 inch wafer shows
        five grating chips. The
        gratings have different curvatures which are specific to the grating
        interferometer geometry. 
        The \ac{SEM} image shows the gold structures and the interrupting bridges
        that prevent the lamellae from collapsing\cite{Kenntner2010}.}
        \label{Fig:grating_mask}
\end{figure}

Due to the high spectral acceptance \cite{Weitkamp2005,Thuering2013c} of the
interferometer ($\SI{50}{\kilo\electronvolt}$ to
more than $\SI{160}{\kilo\electronvolt}$) and the high attenuation efficiencies of
the source and analyzer gratings ($>90\%$ up to
$\SI{160}{\kilo\electronvolt}$), the voltage of the X-ray source was set to
the maximum of $\SI{160}{\kilo\volt}$. With a grating structure height of
approximately $\SI{100}{\micro \metre}$, the field of view in the vertical
direction is limited to one detector pixel row. In the horizontal direction,
the field of view is limited by the grating size and the geometric
magnification of the sample, yielding a maximum field of view of
$\SI{30}{\milli\metre}$. In addition to the standard components (source,
camera, interferometer), two optical slits, one in front of the source
grating, the other in front of the camera, were required for the collimation
of the beam in the vertical direction.

Fig.~\ref{Fig:img_chip} shows a radiographic scan of an electronic chip.
Several resistors and an integrated circuits are located on different layers
on the chip. The images were acquired in scanning mode, using a step size of
$\SI{100}{\micro \metre}$ along the $y$ axis. For a better comparison of the
magnified phase and attenuation images, the attenuation image has been
replaced with the differential attenuation image, which was obtained by
digital differentiation. In the attenuation image, the contrast of the soldering points of
the integrated circuit is reduced underneath the resistors, while in
the phase image, they can clearly be identified. The reduced contrast of the 
soldering points in the absorption image is due to beam hardening. The spectrum
impinging on these soldering point is hardened by the resistors in the upper
layer, resuluting in lower absorption contrast. Due to the weaker energy
dependency of phase shifts ($1/E$ compared to $1/E^3$), phase-contrast images
are less sensitive to beam hardening \cite{Chabior2011a}, which explains the
lower contrast reduction of the soldering points underneath the resistors in the
phase image of the chip. This result shows the benefit of the phase nature in
high energy X-ray imaging, which may be useful to identify flaws in multilayered
structures such as electronic chips.
\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{figures/img_chip_dabs.eps}
    \caption{Radiographic scan of an electronic chip. The image was acquired
        with 24 phase steps per line and an exposure time of \SI{15}{\second} per
    step. The top right image shows the differential absorption image.}
    \label{Fig:img_chip}
\end{figure}

Edge-on illuminated grating interferometry breaks the current limitations of
phase-contrast imaging to the lower diagnostic energy range. Compact
geometries  using conventional X-ray sources for design energies well above
$\SI{100}{\kilo\electronvolt}$ could be realized with the approach. This
enables the examination of materials of higher density or thickness which
would be opaque at lower energies. Finally, the approach is not
limited to X-ray imaging, but could be easily applied to other grating-based
imaging modalities (e.g., with neutrons \cite{Grunzweig2008}) where high
aspect ratios are required.

@@

In Fig.~\ref{Fig:img_chip}, image acquisition involved 24 phase steps per
line and an exposure time of 15 seconds per step. The long
exposure times are mostly constrained by the low average visibility of the
gratings (\SI{5}{\percent}). The exposure time can be shortened with better
gratings and an efficient detector.

@@

\bibitem [{\citenamefont {Chabior}\ \emph {et~al.}(2011)\citenamefont
  {Chabior}, \citenamefont {Donath}, \citenamefont {David}, \citenamefont
  {Bunk}, \citenamefont {Schuster}, \citenamefont {Schroer},\ and\
  \citenamefont {Pfeiffer}}]{Chabior2011a}%
  {Chabior}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Donath}},
  {\bibfnamefont {M.}~\bibnamefont {Schuster}}, \bibinfo {author}
  {\bibfnamefont {C.}~\bibnamefont {Schroer}}, \ and\ \bibinfo {author}
  {\bibfnamefont {F.}~\bibnamefont {Pfeiffer}},\ }\href {\doibase
  10.1118/1.3553408} {\bibfield  {journal} {\bibinfo  {journal} {Medical
  Physics}\ }\textbf {\bibinfo {volume} {38}},\ \bibinfo {pages} {1189}
  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
